---
author:
    - "Pol Sole-Navais"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    html_document:
        theme: default
        highlight: tango
        number_sections: no
        toc: yes
        toc_depth: 3
        toc_float:
            collapsed: no
            smooth_scroll: yes
---

```{r echo=F, include= F}
cohort= unlist(strsplit(snakemake@input[[1]], '/'))[9]
pheno= unlist(strsplit(snakemake@input[[1]], '/'))[8]
```

---
title: "`r paste0('File-level QC: ', cohort, ' and phenotype: ', pheno)`"
---



```{r include=FALSE, echo=F}
library("ggplot2")
library("dplyr")
library("knitr")
library("tidyr")
library(cowplot)
library("kableExtra")
library("data.table")
options(warn=-1)
#opts_chunk$set(dpi=300, out.width="300px")

```

# Summary stats  


```{r echo=F}

d= fread(snakemake@input[[1]], h=T)

```

This report contains a general and automated QC for `r cohort`, phenotype: `r pheno`. A total of `r nrow(d)` genetic variants have passed an initial QC.

```{r echo=F}
kable(summary(select(d, BETA, SE, pvalue, EAF, N)))

```



# Frequency plot against HRC 


```{r echo=F}

#dec= fread(snakemake@input[[3]],h=T, select= c('CHR', 'POS', 'BETA', 'SE', 'EFF', 'REF', 'EAF', 'pvalue'))

#names(dec)= c('CHR', 'POS', 'BETA_dec', 'SE_dec', 'A1_dec', 'A2_dec', 'EAF_dec', 'pvalue_dec')

#dec$BETA_dec= ifelse(dec$A1_dec> dec$A2_dec, dec$BETA_dec, -1* dec$BETA_dec)
#dec$EAF_dec= ifelse(dec$A1_dec> dec$A2_dec, dec$EAF_dec, 1- dec$EAF_dec)

x= fread(snakemake@input[[2]], h=T)


df= d
df$CHR= as.numeric(df$CHR)
x$CHR= as.numeric(x$CHR)
df$EAF= ifelse(df$EFF> df$REF, df$EAF, 1 - df$EAF)

df= inner_join(df, x, on= c('CHR', 'POS'))

df= filter(df, (EFF== ea & REF== oa) | (REF== ea & EFF== oa))

if (nrow(df)>0){
ggplot(df, aes(eaf, EAF)) +
geom_point(alpha= 1/10) +
theme_cowplot(12) +
xlab('EAF HRC') +
ylab('EAF Sample')
} else{
print('No match')
}

rm(df)
rm(x)

```

# Against DECODE  

```{r echo=F}



d$BETA= ifelse(d$REF> d$EFF, d$BETA, -1* d$BETA)
#dec= inner_join(d, dec, on= c('CHR', 'POS'))

#dec= filter(dec, EFF== A1_dec, REF== A2_dec, pvalue_dec< 0.001)

#dec= group_by(dec, CHR) %>% slice_sample(n= 1000, replace= T)

#ggplot(dec, aes(BETA_dec/SE_dec, BETA/SE)) +
#geom_point(alpha= 1/10) +
#theme_cowplot(12) +
#xlab('DECODE z-score') +
#ylab('Observed z-score')

#rm(dec)

```

# Z-Pvalue plot  


```{r echo= F}
d$exp_pvalue= pchisq((d$BETA/d$SE)^2, df=1, lower=F)

ggplot(group_by(d, CHR) %>% sample_n(10000, replace=T ), aes(-log10(exp_pvalue), -log10(pvalue))) +
geom_point(alpha= 1/10) +
theme_cowplot() +
xlab('Expected pvalue') +
ylab('Observed pvalue')

```

Out of `r nrow(d)` genetic variants, `r sum(d$pvalue != d$exp_pvalue)` (`r sum(d$pvalue != d$exp_pvalue) /nrow(d) `) are not equal to the expected pvalue for the z-score.

# QQplot  

```{r echo=F}

df= arrange(d, pvalue) %>% mutate(exp1= -log10(1:length (pvalue)/length (pvalue)))

ggplot(filter(df, pvalue<0.05), aes(exp1, -log10(pvalue))) +
  geom_point(size= 0.4) +
  geom_abline(intercept = 0, slope = 1, alpha = .5) +
theme_cowplot(12, font_size= 12) +
xlab('Expected (-log10(p-value))') +
ylab('Observed (-log10(p-value))')

```


